Reduced diversity and increased specialization of Alaska salmon fishermen:

### 2016-10-24 salmon/paper/salmon_paper.Rmd
typos

Understanding the effects of species diversification on salmon fishers is particularly important, as the number of salmon specialists (defined as fishers holding a single salmon permit) has increased over the last 30 years as a fraction of the total [@anderson_benefits]. For these limited entry salmon fisheries, general trends that have emerged are that more people are participating since the early 2000s (Fig. A1), median revenue is increasing (across individuals, Fig. A2), and landed species diversity declining over time (averaged across indivduals, Fig. A6). Across different salmon fisheries, other than a slight negative effect of diversification on variability (similar to [@anderson_benefits]), there appears to be little correlation between diversity and revenue (Fig. A4). In terrestrial agriculture, there are numerous examples where diversification may affect the mean or CV but not both [@purdy_financial_1997], and the same result may be true for salmon fishers. Further, these impacts of diversification are variable over time.

@@

In addition to the changing species composition of fish avaialble to be harvested, the benefits (or costs) of diversification are affected by a number of exogenous drivers, including marketing, prices, new processing technology [@knapp_trends]. [Break out as own paragraph+add more examples?] Following price crashes in the early 2000s, many of these factors have been positive for salmon fishermen (increasing trends in prices and catches, Fig. A3, A12). In contrast, the late 1980s - early 1990s experienced high abundance and low prices (in part due to collapses of foreign currencies). Combined with the increased trend in hatchery production and resulting changes in species compositon of adult salmon, these external factors may have also altered incentives for diversification. If there's an incentive to substitute high value species (Chinook, sockeye) with more abundant low value species (pink, chum), the analog in the terrestrial framring world would be crop substitution of low-volume high value crops (spices) with high-volume low value crops (cereals) (other examples in literature = cocoa, tobacco).

@@

Hatchery release trends for some regions included in our data have been summarized previously (e.g. [@ward_evaluating] summarized the trends for Prince William Sound). Release patterns have been dominated by pink and chum salmon (in recent years 2012-2014 in PWS, for example, pink and chum salmon represent 80% and 18% of total releases, [@ward_evaluating]) and production has increased since the late 1970s (217 million juvenile pink salmon in PWS in 1985 versus 672 million in 2014). Though our analysis with CFEC data doesn't separate hatchery versus wild fish, as a contrast we can make coarse comparisons across regions and species to examine potential effects of hatchery fish on variability. For example, we can compare the variation in the largely hatchery pink salmon returns (earnings and weight) from Prince William Sound to the largely wild pink salmon returns from Southeast Alaska. Similarly, we can compare the variability in earnings from species dominated by hatchery returns (pink, chum) to those with less hatchery influence (Chinook, sockeye). To make these comparisons, we aggregated the earnings and landings by species and permit, then calculated 3-year rolling CVs. As a second comparison, we examined whether long-term participants in any of the fisheries had experienced a change in variability. Similar to [@kasperski_income_2013] who calculated CVs of vessels before and after significant managment events, we identified individuals who fished the same permit for at least three years in the first decade of our data (1985-1995) and at least three years in the last decade of the data (2004-2014).

@@

To ask how benefits of diversification may have changed over time, we extend the model of [@anderson_benefits] to also include time-varying random effects in the slope coefficients of diversification. We define the variables $\Delta { S }_{ t }=\log\left( { S }_{ t }/{ S }_{ t-1 } \right)$ to represent the percent change in species diversity of the catch between years and $\Delta { D }_{ t }=\log\left( { D }_{ t }/{ D }_{ t-1 } \right)$ to represent the percent change in fishing effort (days fished) between years. The model for the expected log revenue of individual *i* in year *t*, $$\widehat { \log\left( { R }_{ i,t } \right)  } ={ B }_{ 0,j,t }+{ B }_{ 1,j,t }\cdot \Delta S_{ i,t }+{ B }_{ 2,j,t }\cdot \left( \Delta S_{ i,t }\cdot { S }_{ i,t-1 } \right) +{ B }_{ 3,j }\cdot \Delta D_{ i,t }+\log\left( { R }_{ i,t-1 } \right) $$, where *j* subscripts the salmon permit (Table 1). We model the time-varying permit specific intercepts as random and exchangeable across years, ${ B }_{ 0,j,t }\sim \mathrm{Normal}\left( { u }_{ { B }_{ 0 } },{ \sigma  }_{ { B }_{ 0 } } \right)$. The term ${ B }_{ 1,j,t }$ represents the permit level random effect of diversifying species. Our model extends that of [@anderson_benefits] to include the interaction terms $\Delta S_{ t }\cdot { S }_{ t-1 }$ which allow the effect of diversification to be dependent on the previous level of diversification. We assume this relationship is linear, and the coefficients ${ B }_{ 1,j,t }$ and ${ B }_{ 2,j,t }$ allow this to vary by permit and through time. The flexibility in these terms allows the benefit of increasing species diversity to be positive in some years (generalizing benefits revenue), negative in other years (specializing benefits revenue) and neutral in other years (revenue is independent of alterations to catch composition). We constrain the random effects associated with diversification to be modeled as autoregressive processes with shared variances across strategies, for example: ${ B }_{ 1,j,t }\sim \mathrm{Normal}\left( { B }_{ 1,j,t-1 },{ \sigma  }_{ { B }_{ 1 } } \right)$. Like [@anderson_benefits] we also allow changing effort (measured as percent change in days fished) ${ B }_{ 4,j }$ to vary by strategy, but we do not allow this effect to vary through time. Finally, the term $\log\left( { R }_{ i,t-1 } \right)$ is included as an offset [@anderson_benefits]. 
  Like [@anderson_benefits], we also included all of the above covariates in the variance model. We assumed the residuals from the mean model described above to be normally distributed, $$\log\left( { R }_{ i,t } \right) \sim \mathrm{Normal}\left( \widehat { \log\left( { R }_{ i,t } \right)  } ,\widehat { { \sigma  }_{ i,t } }  \right) $$. While conventional hierarchcial linear regression assumes variances to be constant or shared by grouping variables, we also included fixed and random covariates, making the model akin to variance function regression or stochastic volatility models (CITE). Using the same notation as for the mean, we can express the log standard deviation as $$\widehat { \log\left( { \sigma  }_{ i,t } \right)  } ={ G }_{ 0,j }+{ G }_{ 1,j,t }\cdot \Delta S_{ i,t }+{ G }_{ 2,j,t }\cdot \left( \Delta S_{ i,t }\cdot { S }_{ i,t-1 } \right) +G_{ 3,j }\cdot \Delta D_{ i,t }$$. Unlike the linear model for the mean, we did not include an offset term. Further, we constrained the variance intercepts ${ G }_{ 0,j }$ to be temporally constant by strategy, instead of the dynamic nature of the intercepts in the mean model. Like the mean model, the random effects were constrained to be autoregressive, ${ G }_{ 1,j,t }\sim \mathrm{Normal}\left( { G }_{ 1,j,t-1 },{ \sigma  }_{ { G }_{ 1 } } \right)$, and covariates associated with days fished and species diversity were included as predictors.

@@

Figure 3 shows that one component of variability is the year-to-year variation that individuals experience as a result of external forces (stock abundance, prices, etc) as wekk as common fluctuations in revenue experienced by all permit holders (Anderson et al., n.d.). This shows the time varying intercepts on the mean -- negative equals losing revenue on average, while positive implies increasing revenue. The panels of this figure show big differences between permit types (gillnets tend to have lower year-to-year variation). Some permits have extreme variation; for example, S01E (PWS seine) has fluctuations between exp(-1) and exp(+1), corresponding to year-to-year changes of 36-270%.

Figure 4 shows the time varying nature of of the model for all permits. These are the time series of benefits of changing from effective species diversity of 1 to 2, and here we are isolating year-to-year differences in revenue (the intercepts) from diversification benefits. In some cases the effect is positive (fishing more spp yields more revenue). But in many cases this is not the case, and the trend is negative (indicating that specialization is more profitable over time). In some cases, these time series mirror the aggregate diversity in the raw time series (Fig. 1). I think we also expect this result - fishermen increase diversity or not, because it's profitable.

Figure 5 shows that temporal trends, by permit, in going from 1 to 2 species fished on the variance (again removing the overall intercepts). From [@anderson_benefits] we'd expect diversification to redudce variance. A few examples of how to read this: S01A is slightly positve for most of the time series, but rapidly increases in the last 10 years -- this means generalists with diverse portfolios have had increases in year-to-year variability. For S01E, the start and end of the time series are associated with these 'costs' to diversification. For S03E, we see a more gradual trend (cost of diversification increases).

>>>>

Understanding the effects of species diversification on salmon fishers is particularly important, as the number of salmon specialists (defined as fishers holding a single salmon permit) has increased over the last 30 years as a fraction of the total [@anderson_benefits]. For these limited entry salmon fisheries, general trends that have emerged are that more people are participating since the early 2000s (Fig. A1), median revenue is increasing (across individuals, Fig. A2), and landed species diversity declining over time (averaged across individuals, Fig. A6). Across different salmon fisheries, other than a slight negative effect of diversification on variability (similar to [@anderson_benefits]), there appears to be little correlation between diversity and revenue (Fig. A4). In terrestrial agriculture, there are numerous examples where diversification may affect the mean or CV but not both [@purdy_financial_1997], and the same result may be true for salmon fishers. Further, these impacts of diversification are variable over time.

@@

In addition to the changing species composition of fish available to be harvested, the benefits (or costs) of diversification are affected by a number of exogenous drivers, including marketing, prices, new processing technology [@knapp_trends]. [Break out as own paragraph+add more examples?] Following price crashes in the early 2000s, many of these factors have been positive for salmon fishermen (increasing trends in prices and catches, Fig. A3, A12). In contrast, the late 1980s - early 1990s experienced high abundance and low prices (in part due to collapses of foreign currencies). Combined with the increased trend in hatchery production and resulting changes in species composition of adult salmon, these external factors may have also altered incentives for diversification. If there's an incentive to substitute high value species (Chinook, sockeye) with more abundant low value species (pink, chum), the analog in the terrestrial farming world would be crop substitution of low-volume high value crops (spices) with high-volume low value crops (cereals) (other examples in literature = cocoa, tobacco).

@@

Hatchery release trends for some regions included in our data have been summarized previously (e.g. [@ward_evaluating] summarized the trends for Prince William Sound). Release patterns have been dominated by pink and chum salmon (in recent years 2012-2014 in PWS, for example, pink and chum salmon represent 80% and 18% of total releases, [@ward_evaluating]) and production has increased since the late 1970s (217 million juvenile pink salmon in PWS in 1985 versus 672 million in 2014). Though our analysis with CFEC data doesn't separate hatchery versus wild fish, as a contrast we can make coarse comparisons across regions and species to examine potential effects of hatchery fish on variability. For example, we can compare the variation in the largely hatchery pink salmon returns (earnings and weight) from Prince William Sound to the largely wild pink salmon returns from Southeast Alaska. Similarly, we can compare the variability in earnings from species dominated by hatchery returns (pink, chum) to those with less hatchery influence (Chinook, sockeye). To make these comparisons, we aggregated the earnings and landings by species and permit, then calculated 3-year rolling CVs. As a second comparison, we examined whether long-term participants in any of the fisheries had experienced a change in variability. Similar to [@kasperski_income_2013] who calculated CVs of vessels before and after significant management events, we identified individuals who fished the same permit for at least three years in the first decade of our data (1985-1995) and at least three years in the last decade of the data (2004-2014).

@@

To ask how benefits of diversification may have changed over time, we extend the model of [@anderson_benefits] to also include time-varying random effects in the slope coefficients of diversification. We define the variables $\Delta { S }_{ t }=\log\left( { S }_{ t }/{ S }_{ t-1 } \right)$ to represent the percent change in species diversity of the catch between years and $\Delta { D }_{ t }=\log\left( { D }_{ t }/{ D }_{ t-1 } \right)$ to represent the percent change in fishing effort (days fished) between years. The model for the expected log revenue of individual *i* in year *t*, $$\widehat { \log\left( { R }_{ i,t } \right)  } ={ B }_{ 0,j,t }+{ B }_{ 1,j,t }\cdot \Delta S_{ i,t }+{ B }_{ 2,j,t }\cdot \left( \Delta S_{ i,t }\cdot { S }_{ i,t-1 } \right) +{ B }_{ 3,j }\cdot \Delta D_{ i,t }+\log\left( { R }_{ i,t-1 } \right) $$, where *j* subscripts the salmon permit (Table 1). We model the time-varying permit specific intercepts as random and exchangeable across years, ${ B }_{ 0,j,t }\sim \mathrm{Normal}\left( { u }_{ { B }_{ 0 } },{ \sigma  }_{ { B }_{ 0 } } \right)$. The term ${ B }_{ 1,j,t }$ represents the permit level random effect of diversifying species. Our model extends that of [@anderson_benefits] to include the interaction terms $\Delta S_{ t }\cdot { S }_{ t-1 }$ which allow the effect of diversification to be dependent on the previous level of diversification. We assume this relationship is linear, and the coefficients ${ B }_{ 1,j,t }$ and ${ B }_{ 2,j,t }$ allow this to vary by permit and through time. The flexibility in these terms allows the benefit of increasing species diversity to be positive in some years (generalizing benefits revenue), negative in other years (specializing benefits revenue) and neutral in other years (revenue is independent of alterations to catch composition). We constrain the random effects associated with diversification to be modeled as autoregressive processes with shared variances across strategies, for example: ${ B }_{ 1,j,t }\sim \mathrm{Normal}\left( { B }_{ 1,j,t-1 },{ \sigma  }_{ { B }_{ 1 } } \right)$. Like [@anderson_benefits] we also allow changing effort (measured as percent change in days fished) ${ B }_{ 4,j }$ to vary by strategy, but we do not allow this effect to vary through time. Finally, the term $\log\left( { R }_{ i,t-1 } \right)$ is included as an offset [@anderson_benefits]. 
  Like [@anderson_benefits], we also included all of the above covariates in the variance model. We assumed the residuals from the mean model described above to be normally distributed, $$\log\left( { R }_{ i,t } \right) \sim \mathrm{Normal}\left( \widehat { \log\left( { R }_{ i,t } \right)  } ,\widehat { { \sigma  }_{ i,t } }  \right) $$. While conventional hierarchical linear regression assumes variances to be constant or shared by grouping variables, we also included fixed and random covariates, making the model akin to variance function regression or stochastic volatility models (CITE). Using the same notation as for the mean, we can express the log standard deviation as $$\widehat { \log\left( { \sigma  }_{ i,t } \right)  } ={ G }_{ 0,j }+{ G }_{ 1,j,t }\cdot \Delta S_{ i,t }+{ G }_{ 2,j,t }\cdot \left( \Delta S_{ i,t }\cdot { S }_{ i,t-1 } \right) +G_{ 3,j }\cdot \Delta D_{ i,t }$$. Unlike the linear model for the mean, we did not include an offset term. Further, we constrained the variance intercepts ${ G }_{ 0,j }$ to be temporally constant by strategy, instead of the dynamic nature of the intercepts in the mean model. Like the mean model, the random effects were constrained to be autoregressive, ${ G }_{ 1,j,t }\sim \mathrm{Normal}\left( { G }_{ 1,j,t-1 },{ \sigma  }_{ { G }_{ 1 } } \right)$, and covariates associated with days fished and species diversity were included as predictors.

@@

Figure 3 shows that one component of variability is the year-to-year variation that individuals experience as a result of external forces (stock abundance, prices, etc) as well as common fluctuations in revenue experienced by all permit holders (Anderson et al., n.d.). This shows the time varying intercepts on the mean -- negative equals losing revenue on average, while positive implies increasing revenue. The panels of this figure show big differences between permit types (gillnets tend to have lower year-to-year variation). Some permits have extreme variation; for example, S01E (PWS seine) has fluctuations between exp(-1) and exp(+1), corresponding to year-to-year changes of 36-270%.

Figure 4 shows the time varying nature of the model for all permits. These are the time series of benefits of changing from effective species diversity of 1 to 2, and here we are isolating year-to-year differences in revenue (the intercepts) from diversification benefits. In some cases the effect is positive (fishing more spp yields more revenue). But in many cases this is not the case, and the trend is negative (indicating that specialization is more profitable over time). In some cases, these time series mirror the aggregate diversity in the raw time series (Fig. 1). I think we also expect this result - fishermen increase diversity or not, because it's profitable.

Figure 5 shows that temporal trends, by permit, in going from 1 to 2 species fished on the variance (again removing the overall intercepts). From [@anderson_benefits] we'd expect diversification to reduce variance. A few examples of how to read this: S01A is slightly positive for most of the time series, but rapidly increases in the last 10 years -- this means generalists with diverse portfolios have had increases in year-to-year variability. For S01E, the start and end of the time series are associated with these 'costs' to diversification. For S03E, we see a more gradual trend (cost of diversification increases).
